Manuscript Te example

### 2020-11-10 ManuscriptTe.Rmd
Update ms

---
title: "Manuscript Te example"
author: "Lauren Buckley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

@@

```{r}
##Load TrenchR package
library(TrenchR)
library(maptools)
library(ggplot2)
```

#Microclimate calculations
```{r}
setwd("/Volumes/GoogleDrive/Shared drives/TrEnCh/Projects/TrenchR/data/")
C1all=read.csv("microclimate_C1.csv", sep=",", na.strings = c("-49.99","NA"))
```

```{r}
#ANALYZE SURFACE ROUGHNESS AND PROFILES
C1all$day=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$mday
C1all$hour=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$hour

@@

We can estimate surface roughness across hours:
```{r}
#surface roughness
z0= apply(C1.hr[, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], FUN=surface_roughness, MARGIN=1, zr=c(0.25,0.5,1.0,1.5))

@@

We calculate the temperature and windspeed profiles for a given hour using multiple functions in TrenchR.
```{r}
z.seq <- seq(0, 2, by = 0.1)

@@

u.seq2= sapply(z.seq, FUN=wind_speed_profile, u_r=C1.sub[hr,"Wind0.5m"], zr=0.5, z0=z0[hr])

@@

We then plot the temperature and windspeed profiles.
```{r}
plot(t.seq1, z.seq, type = "l", xlab = "Temperature (C)", ylab = "z (height above ground in m)")
points(t.seq2, z.seq, type = "l", lty="dashed")
points(t.seq3, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Air0.25m","Air0.5m","Air1.0m","Air1.5m")], c(0.25,0.5,1,1.5), type = "l", col="red")

plot(u.seq1, z.seq, type = "l", xlab = "windspeed (m/s)", ylab = "z (height above ground in m)")
points(u.seq2, z.seq, type = "l", lty="dashed")
points(u.seq3, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], c(0.25,0.5,1,1.5), type = "l", col="red")
```

#Te calculation
```{r}

#Grasshopper data
setwd("/Volumes/GoogleDrive/Shared drives/TrEnCh/Projects/TrenchR/data/")
spec.dat=read.csv("SpecData.csv")

@@

#ditch problematic temperatures
dat$Temp[dat$Temp< -10]=NA

@@

#----------------

z<-0.001 #specify distance from ground

@@

#Subset to August
dat.dt<- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST") #as.Date
dataug= subset(dat, dat.dt$mon==7)
#subset to day
dat= subset(dataug, dataug$Rad>5)

@@

```{r}
##Calculate Te for each species
dat$Te<-Tb_grasshopper(T_a=dat$Tgrass, T_g=dat$SoilTemp, u=dat$Wind, H=dat$Rad, K_t=0.7, psi=dat$psi, L=0.0211, Acondfact=0.01)  #T_a=dat$Tgrass
#fix infinite
dat$Te[is.infinite(dat$Te)]<-NA
#ditch big numbers for now
dat$Te[dat$Te>100]<-NA
```

```{r}
#estimate convective heat transfer coefficient

#Estimate using empirical relationship
dat$H_L=heat_transfer_coefficient(V=dat$Wind,D=0.0211,K=0.025, nu= 15.3 * 10^(-6) , taxa="cylinder")

@@

```{r}
#Gates calculation
TeGates= rep(NA, nrow(dat) )

@@

```{r}
#aggregate by day and hour
dat.dh= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour, dat$J), FUN=mean, na.rm=TRUE))
```

```{r}
#aggregate by hour
dat.h= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour), FUN=mean, na.rm=TRUE))
```

```{r}
library(reshape2)
#convert to long format
dat.s= melt(dat.h, id = c("Elev", "J","hour","psi"))

@@

plot.t=ggplot(data=dat.s, aes(x=Jhr, y = value, color=variable))+geom_line(aes(lty=variable) )+facet_grid(Elev~1)
```

@@

pdf("TeTime.pdf", height = 10, width = 15)
plot.t
dev.off()

>>>>

---
title: "Manuscript Te example"
author: "Lauren Buckley"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
  rmarkdown::html_vignette: default
---

@@

```{r}
##Load TrenchR package
library(TrenchR)
library(maptools)
library(ggplot2)
library(reshape)
```

#APPENDIX S1
We provide an additional example using microclimate data measured along an elevation gradient in CO, USA to illustrate use of TrenchR. We first use a times series of air temperatures and wind speeds collected at multiple heights to examine profiles and surface roughness. We then use a time series of air and surface temperature, windspeed, and solar radiation collected a a single height to implement and test a biophysical model for grasshoppers.

#Microclimate models
```{r}
setwd("/Volumes/GoogleDrive/Shared drives/TrEnCh/Projects/TrenchR/data/")
C1all=read.csv("microclimate_C1.csv", sep=",", na.strings = c("-49.99","NA"))
```

We first examine temperature and wind speed profiles and use them to estimate surface roughness.
```{r}
#ANALYZE SURFACE ROUGHNESS AND PROFILES
C1all$day=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$mday
C1all$hour=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$hour

@@

We can use wind speed variation across elevations to estimate surface roughness across hours.
```{r}
#surface roughness
z0= apply(C1.hr[, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], FUN=surface_roughness, MARGIN=1, zr=c(0.25,0.5,1.0,1.5))

@@

We calculate the temperature and wind speed profiles for a given hour using multiple functions in TrenchR.
```{r}
z.seq <- seq(0, 2, by = 0.1)

@@

#u.seq2= sapply(z.seq, FUN=wind_speed_profile, u_r=C1.sub[hr,"Wind0.5m"], zr=0.5, z0=z0[hr])

@@

We then plot the temperature and wind speed profiles.
```{r}
plot(t.seq1, z.seq, type = "l", xlab = "Temperature (C)", ylab = "z (height above ground in m)")
points(t.seq2, z.seq, type = "l", lty="dashed")
points(t.seq3, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Air0.25m","Air0.5m","Air1.0m","Air1.5m")], c(0.25,0.5,1,1.5), type = "p", col="red")

plot(u.seq1, z.seq, type = "l", xlab = "windspeed (m/s)", ylab = "z (height above ground in m)")
#points(u.seq2, z.seq, type = "l", lty="dashed")
points(u.seq3, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], c(0.25,0.5,1,1.5), type = "p", col="red")
```

#Biophysical models
We first read and process time series of environmental data.
```{r}
#Grasshopper data
setwd("/Volumes/GoogleDrive/Shared drives/TrEnCh/Projects/TrenchR/data/")
spec.dat=read.csv("SpecData.csv")

@@

#omit problematic temperatures
dat$Temp[dat$Temp< -10]=NA

@@

@@

#Subset to August
dat.dt<- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST") #as.Date
dataug= subset(dat, dat.dt$mon==7)
#subset to day
dat= subset(dataug, dataug$Rad>5)
```

We scale the air temperature data from 0.5m to grasshopper height (0.001m).
```{r}
z<-0.001 #specify distance from ground

@@

We then use biophysical models to estimate body temperatures based on air and surface temperatures, wind speed, and solar radiation. We first use a biophysical model specifically configured for grasshoppers.
```{r}
##Calculate Te for each species
dat$Te<-Tb_grasshopper(T_a=dat$Tgrass, T_g=dat$SoilTemp, u=dat$Wind, H=dat$Rad, K_t=0.7, psi=dat$psi, L=0.0211, Acondfact=0.01)
#fix infinite
dat$Te[is.infinite(dat$Te)]<-NA
#ditch big numbers for now
dat$Te[dat$Te>100]<-NA
```

We then implement a generic energy budget. We first illustrate the estimation of convective heat transfer coefficients as a function of wind speed.
```{r}
#Estimate heat transfer coefficient using empirical relationship
dat$H_L=heat_transfer_coefficient(V=dat$Wind,D=0.0211,K=0.025, nu= 15.3 * 10^(-6) , taxa="cylinder")

@@

We then implement the energy budget as follows.
```{r}
#Gates calculation
TeGates= rep(NA, nrow(dat) )

@@

We aggregate the data and plot average conditions across sites.
```{r}
#aggregate by hour
dat.h= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour), FUN=mean, na.rm=TRUE))
#aggregate by day and hour
dat.dh= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour, dat$J), FUN=mean, na.rm=TRUE))

#convert to long format
dat.s= melt(dat.h, id = c("Elev", "J","hour","psi"))

@@

#plot
dat.s= subset(dat.s, dat.s$variable %in% c("Temp","SoilTemp","SanguinipesTemp","Te","Tgrass") )
#,"TeGates"

plot.t= ggplot(data=dat.s, aes(x=Jhr, y = value, color=variable))+geom_line(aes(lty=variable))+facet_grid(Elev~1)
```

@@

pdf("TeTime2.pdf", height = 10, width = 15)
plot.t
dev.off()
